---
title: Getting Started
output:
  revealjs::revealjs_presentation:
    center: no
    css: ../css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: false
    theme: solarized
    transition: slide
---

```{r include=FALSE}
source('common.R')
```

##  Options

* Don't save history, or .Rdata
* Panes
* Editor color
  
## Other options

* Code diagnostics
* Change panel positions
* Consider font and color
* Autocompletion
    
## Script

* Create a script
* Save it, open it

## Keyboard shortcuts

* ALT+SHIFT+K

My favories:

* ALT+-
* CTRL+ENTER
* CTRL+1
* CTRL+2
* CTRL+SHIFT+M
* CTRL+L
* CTRL+SHIFT+N
* CTRL+W
* CTRL+SHIFT+F
* CTRL+SHIFT+B

## Code execution

CTRL+ENTER

90% of your code writing and execution will be done in scripts. Use the console for getting help or one-off operations like: package installation, light debugging.

## Speaking of debugging

R has a debugging framework. I rarely use it.

The combination of environment pane and REPL does 98% of what I need. YMMV

## The working directory

```{r}
getwd()
```

You will probably never have a good reason to call `setwd()`. Use a project instead.

## Projects

* Collect all of the files relevant to a project
* Automatically set the working directory 
* Intermediate stuff:
  * Version control
  * Package development
  * CTRL-SHIFT-B
  
Let's all create one now.

## Packages

```{r eval=FALSE}
install.packages('actuar')
library(actuar)
```

Also have a look at the packages pane in RStudio.

## Data in packages

```{r eval=FALSE}
data(package = 'Lahman')
```

## Installation vs. library

1. `install.packages()` copies files to your computer.
2. `library()` finds those files and loads them into memory.
3. When you turn off your computer, the files are unloaded from memory.
4. Files on your computer don't disappear when you turn off your computer.

## Package installation folder

```{r}
.libPaths()
```

Create a "universal" library by creating a file called ".Renviron".

```
R_LIBS="~/Rlibs"
```

https://stat.ethz.ch/R-manual/R-devel/library/base/html/Startup.html

## Exercise

1. Create a universal library. Ensure that it's working by deleting one of the packages and reinstalling.
2. Create an RStudio project to contain your work for this training.