---
title: "Simulation"
output:
  revealjs::revealjs_presentation:
    center: no
    css: ../css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: false
    theme: solarized
    transition: slide
---

```{r include=FALSE}
source('common.R')
```

## Probability distributions

All probability distributions have four basic functions:

* r dist - Random number generation
* d dist - Density function
* p dist - Cumulative distribution
* q dist - Quantiles

## Simulate

```
r_[DIST](n, PARAMS)

rnorm(10, 5, 10)
rlnorm(10)
rexp(10)
rgamma(10)
```

## Simulate

```{r}
dist_mean <- 10e3
dist_cv <- 2.5
dist_sd <- dist_mean * dist_cv

sims <- 5e3
norms <- rnorm(sims, dist_mean, dist_sd)

norms %>% head()
```

## Histogram

```{r}
hist(norms)
```

## Store simulations in a data frame

```{r }
tbl_sims <- data.frame(
    x_exp = rexp(sims, 1 / dist_mean)
  , x_norm = rnorm(sims, dist_mean, dist_sd)
)
```

## Visualize histogram

```{r}
tbl_sims %>% 
  ggplot(aes(x = x_norm)) +
  geom_histogram()
```

## Visualize density

```{r}
tbl_sims %>% 
  ggplot(aes(x_norm)) + 
  geom_density()
```

## Visualize both

```{r}
tbl_sims %>% 
  ggplot() + 
  geom_density(aes(x_exp), alpha = 0.8, fill = 'red') + 
  geom_density(aes(x_norm), alpha = 0.8, fill = 'blue')
```

# Density function

## Density function

```{r}
plot_points <- 500
x_lims <- dist_mean + c(-1, 1) * 3 * dist_sd
tbl_plot <- data.frame(
    x = seq(x_lims[1], x_lims[2], length.out = plot_points)
  )
tbl_plot$density_norm <- dnorm(tbl_plot$x, dist_mean, dist_sd)
```

## Vis

```{r}
tbl_plot %>% 
  ggplot(aes(x, density_norm)) + 
  geom_line()
```

## Distribution

```{r}
tbl_plot$cdf_norm <- pnorm(tbl_plot$x, dist_mean, dist_sd)
tbl_plot %>% 
  ggplot(aes(x, cdf_norm)) + 
  geom_line()
```

## Quantiles

You're pricing treaty XOL and want to cede the top 1% per claim. Where is your attachment?

```{r}
tails <- c(0.95, 0.98, 0.99)
qnorm(tails, dist_mean, dist_sd)
```

## Your turn

Generate density and cdf plots for an exponential distribution


# Compound loss distribution

## Compound loss distribution

```{r }
set.seed(1234)
policy_years <- 2001:2010
freq <- 1e3
num_claims <- rpois(length(policy_years), freq)
dist_means <- dist_mean * 1.05 ^ (policy_years - min(policy_years))

severity <- mapply(rnorm, num_claims, dist_means, MoreArgs = list(sd = dist_sd))
tbl_claim <- data.frame(
    policy_year = rep(policy_years, num_claims)
  , severity = unlist(severity))
```

## Visualize

```{r}
tbl_claim %>% 
  ggplot(aes(severity)) + 
  geom_density(alpha = 0.6)
```

```{r}
tbl_claim %>% 
  ggplot(aes(severity, fill = as.factor(policy_year))) + 
  geom_density(alpha = 0.6)
```

## Visualize

```{r}
tbl_claim %>% 
  ggplot(aes(severity, fill = as.factor(policy_year))) + 
  geom_density() + 
  facet_wrap(~ policy_year)
```

## 

```{r}
tbl_claim %>% 
  group_by(policy_year) %>% 
  summarise(med = median(severity))
```


# Misc

## `sample`

Generate a random sample of any discrete set of values. 

```{r }
set.seed(1234)
sample(1:100, 10)
```

Use the `prob` argument to weight the probabilities.

```{r }
sample(1:3, prob = c(1, 1, 100), replace = TRUE)
```

## 

To randomize the order of a vector, leave the default value for replace=FALSE.

```{r }
set.seed(1234)
letters[sample(length(letters))]
```

<!-- ## Questions -->

<!-- * Draw a lognormal distribution with a mean of $10,000 and a CV of 30%. -->
<!-- * For that distribution, what is the probability of seeing a claim greater than $100,000? -->
<!-- * Generate 100 and 1,000 observations from that distribution. -->
<!-- * Draw a histogram for each sample. -->
<!-- * What are the mean, standard deviation and CV of each sample? -->

<!-- ## Answers -->

<!-- ```{r } -->
<!-- severity <- 10000 -->
<!-- CV <- .3 -->
<!-- sigma <- sqrt(log(1 + CV^2)) -->
<!-- mu <- log(severity) - sigma^2/2 -->
<!-- plot(function(x) dlnorm(x), mu, sigma, ylab="LN f(x)") -->
<!-- ``` -->

<!-- ## -->

<!-- ```{r } -->
<!-- set.seed(1234) -->
<!-- claims = rlnorm(100, meanlog=log(30000), sdlog=1) -->
<!-- hist(claims, breaks=seq(1, 500000, length.out=40)) -->
<!-- ``` -->