---
title: "Visualization"
output:
  revealjs::revealjs_presentation:
    center: no
    css: ../css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: false
    theme: solarized
    transition: slide
---

```{r include=FALSE}
source('common.R')
```

# Base plotting

## Base plotting

```{r}
library(raw)
data("RegionExperience")

plot(RegionExperience$PolicyYear, RegionExperience$NumClaims)
```

## Add labels

```{r}
plot(
  RegionExperience$PolicyYear
  , RegionExperience$NumClaims
  , pch = 19
  , xlab = 'Policy Year'
  , ylab = '# of claims')
```

## Base plotting

* Kinda ugly
* Customization is possible, but tedious

# ggplot

## ggplot2

gg =  "grammar of graphics" based on Leland Wilkinson

1. Data
2. Mapping
2. Geometric layers
3. Scales

# Data

## Data

```{r error = TRUE}
library(ggplot2)
library(dplyr)
plt_base <- ggplot(RegionExperience)

plt_base

data(COTOR2)
ggplot(COTOR2)
```

## Data

Data must use a data frame.

```{r}
ggplot(1:10)
```

# Mapping

## Mapping

Mappings bind data to visual elements. Every data-derived visual element must be mapped to data.

```{r }
plt_base <- plt_base + aes(x = PolicyYear, y = NumClaims)

plt_base <- ggplot(RegionExperience, aes(PolicyYear, NumClaims))
```

# Layers

## Adding layers

Although we have data and we've mapped to elements of a coordinate system, we haven't specified what the visual elements should be. The `geom_*` family of functions add geometric shapes. 

```{r }
plt_base <- plt_base + geom_point()
plt_base
```

## One step

```{r }
RegionExperience %>% 
  ggplot(aes(x = PolicyYear, y = NumPolicies)) + 
  geom_point()
```

## Try multiple layers

```{r }
RegionExperience %>% 
  ggplot(aes(x = PolicyYear, y = NumPolicies)) + 
  geom_point() + 
  geom_line()
```

## Try that again

```{r }
RegionExperience %>% 
  ggplot(aes(x = PolicyYear, y = NumPolicies)) + 
  geom_point() + 
  geom_line(aes(color = Region))
```

## ggplot vs geom_X

```
ggplot(data, mapping)
geom_X(mapping, data)
```

## Statistical smoothers

```{r}
RegionExperience %>% 
  ggplot(aes(x = PolicyYear, y = NumPolicies)) + 
  geom_point() + 
  geom_smooth(method = lm, aes(color = Region))
```

# Scales

## Store a basic plot in an object

```{r}
plt <- RegionExperience %>% 
  ggplot(aes(x = PolicyYear, y = NumClaims)) + 
  geom_point()
```

## Change the y scale

```{r}
plt + scale_y_continuous(labels = scales::comma)
```

## Change the x breaks

```{r}
plt + scale_x_continuous(breaks = 2001:2010)
```

## Change the x breaks

```{r}
my_breaks <- function(lims) {
  seq(ceiling(lims[1]), floor(lims[2]), by = 1)
}
plt + scale_x_continuous(breaks = my_breaks)
```

# Facets

## Facets

Facets split the data into groups and draws a different plot for each group.

```{r}
ggplot(RegionExperience, aes(x = PolicyYear, y = NumClaims, color = Region)) + 
  geom_point() + 
  facet_wrap(~ Region)
```

# Non-data visual elements

## Non-data visual elements

* labels
* gridlines
* legend
* axis ticks

## Labels

```{r}
plt + labs(x = 'Policy year', y = '# of claims')
```

## Gridlines

```{r}

```

<!-- ## Your turn -->

<!-- 1. Create a scatter plot for policy year and number of claims -->
<!-- 2. Color each point based on region -->
<!-- 2. Add a linear smoother. Which region is showing the greatest increase in claims? -->
<!-- 4. Form the policy frequency by taking the ratio of claims to policies. Plot this. -->

<!-- Extra credit: -->

<!-- 1. Use the state data to create a time series number of claims. Facet by region. -->

<!-- ## Answers -->

<!-- ```{r} -->
<!-- library(raw) -->
<!-- data("RegionExperience") -->
<!-- plt1 <- ggplot(RegionExperience, aes(x = PolicyYear, y = NumClaims)) + geom_point() -->
<!-- plt1 -->

<!-- plt2 <- plt1 + aes(color = Region) -->
<!-- plt2 -->

<!-- plt3 <- plt2 + stat_smooth(method = "lm") -->
<!-- plt3 -->

<!-- RegionExperience$Frequency <- with(RegionExperience, NumClaims / NumPolicies) -->

<!-- plt4 <- ggplot(RegionExperience, aes(x = PolicyYear, y = Frequency, color = Region)) + geom_point() + geom_line() + stat_smooth(method = lm) -->
<!-- plt4 -->
<!-- ``` -->

<!-- ```{r} -->
<!-- data("StateExperience") -->
<!-- pltExtra <- ggplot(StateExperience, aes(x = PolicyYear, y = NumClaims, color = Postal)) + geom_point() + geom_line() -->
<!-- pltExtra + facet_wrap(~ Region) -->
<!-- ``` -->

## Summary

* `ggplot2` is different at first, but will repay your investment.
* Works very well with grouped data to color/facet points.
* Fine-tuning things like axis labels can be a headache, but will get easier. Yes, Excel makes it easier to add data labels and change colors. `ggplot2` makes it easier to work with data.
